\pagestyle{plain}
\null\vfil
\begin{center}
{\bf Acknowledgements}
\end{center}
\setcounter{page}{0}

We are immensely grateful to Aravind Joshi for supporting this
project. 

The principal authors of the current document are: Tilman Becker,
Christine Doran, Dania Egedi, Beth Ann Hockey, Seth Kulick and
B. Srinivas. We are also grateful to Rajesh Bhatt for editorial help.

The following people have contributed to the development of grammars
in the project: Anne Abeille, Rajesh Bhatt, Kathleen Bishop, Sharon
Cote, Beatrice Daille, Christine Doran, Dania Egedi, Jason Frank,
Caroline Heycock, Beth Ann Hockey, Daniel Karp, Seth Kulick, Young-Suk
Lee, Patrick Martin, Megan Moser, Sabine Petillon, Yves Schabes,
Victoria Tredinnick and Raffaella Zanuttini.

The XTAG system has been developed by: Tilman Becker, Richard
Billington, Andrew Chalnick, Dania Egedi, Devtosh Khare, Albert Lee,
David Magerman, Alex Mallet, Patrick Paroubek, Rich Pito, Gilles
Prigent, Anoop Sarkar, Yves Schabes, B. Srinivas, Yuji Yoshiie and
Martin Zaidel. 

We would also like to thank Michael Hegarty, Lauri Karttunen, Anthony Kroch,
Mitchell Marcus, Martha Palmer, Owen Rambow, Phillip Resnik, Beatrice Santorini
and Mark Steedman.

In addition, Jeff Aaronson, Douglas DeCarlo, Mark-Jason Dominus, Mark Foster,
Gaylord Holder, David Mageman, Ken Noble, Steven Shapiro and Ira Winston have
provided technical support.  Adminstrative support was provided by Carolyn
Elken, Jodi Kerper, Christine Sandy and Trisha Yannuzzi.  

This work was partially supported by ARO grant DAAL03-89-0031, ARPA
grant N00014-90-J-1863, NSF STC grant DIR-8920230, and Ben Franklin
Partnership Program (PA) grant 93S.3078C-6. 

\newpage

