\chapter{Getting Around}

This technical report presents the English XTAG grammar as implemented by the
XTAG Research Group at the University of Pennsylvania.  The technical report is
organized in 4 parts, plus a set of appendices.  Part 1 contains general
information about the XTAG system and some of the underlying mechanisms that
help shape the grammar.  Chapter~\ref{intro-FBLTAG} contains an introduction to
the formalism behind the grammar and parser, while chapter~\ref{overview}
contains information about the entire XTAG system.  Linguists interested solely
in the grammar of the XTAG system may safely skip chapters~\ref{intro-FBLTAG}
and \ref{overview}.  Chapter~\ref{underview} contains information on some of
the linguistic principles that underlie the XTAG grammar, including the
distinction between complements and adjuncts, and how case is handled.

The actual description of the grammar begins with Part 2, and in contained in
the following 3 parts.  Parts 2 and 3 contains information on the verb classes
and the types of trees allowed within the verb classes, respectively, while
Part 4 contains information on trees not included in the verb classes (e.g.
NP's, PP's, various modifiers, etc).  Chapter~\ref{table-intro} of Part 2
contains a table that attempts to provide an overview of the verb classes and
tree types by providing a graphical indication of which tree types are allowed
in which verb classes.  This has been cross-indexed to tree figures shown in
the tech report.  Chapter~\ref{verb-classes} contains an overview of all of the
verb classes in the XTAG grammar.  The rest of Part 2 contains more details on
several of the more interesting verb classes, including ergatives, sentential
subjects, sentential complements, small classes, ditransitives, and it-clefts.

Part 3 contains information on some of the tree types that are available within
the verb classes.  These tree types correspond to what would be transformations
in a movement based approach.  Not all of these types of trees are contained in
all of the verb classes.  The table (previously mentioned) in Part 2 contains a
list of the tree types and indicates which verb classes each occurs in.  

Part 4 moves away from the idea of verb classes, and contains information on
`the rest' of the trees in the grammar.  NP's and determiners are presented in
chapter~\ref{det-comparitives}, while the various modifier trees are presented
in chapter~\ref{modifiers}.  A chapter on auxiliary verbs, which are classed
separate from the verb classes, is found in chapter~\ref{auxiliaries}, while
certain types of conjunction are shown in chapter~\ref{conjunction}.  Other
conjunctions, such as subordinating and discourse conjunction, are considered
tree types, and as such as included in the chapter on adjunct clauses
(section~\ref{sub-conj}).  Sentential complements of NP's and PP's are
discussed in section~\ref{NPA}.

Throughout the technical report, mention is occasionally made of changes or
analyses that we hope to incorporate in the future.  Appendix~\ref{future-work}
details a list of these and other future work.  The appendices also contain
information on some of the nitty gritty details of the XTAG grammar, including
the tree naming conventions (appendix~\ref{tree-naming}), and a comprehensive
list of the features used in the grammar (appendix~{\ref{features}).
Appendix~\ref{evaluation} contains an evaluation of the XTAG grammar,
including comparisons with other wide coverage grammars.


