\chapter{Tree Naming conventions}
\label{tree-naming}

The various trees within the XTAG grammar are named more or less according to
the following tree naming conventions.  Although these naming conventions are
generally followed, there are occasional trees that do not strictly follow
these conventions.

\section{Tree Families}
Tree families are named according to the basic declarative tree structure in
the tree family (see section~\ref{family-trees}), but with a T as the first
character instead of an $\alpha$ or $\beta$.

\section{Trees within tree families}
\label{family-trees}

Each tree begins with either an $\alpha$ (alpha) or a $\beta$ (beta) symbol,
indicating whether it is an initial or auxiliary tree, respectively.  Following
an $\alpha$ or a $\beta$ the name may additionally contain one of:

\begin{description}
\item\begin{tabular}{ll}
I&imperative\\
E&ergative\\
N{0,1,2}&relative clause\{position\}\\
G&NP gerund\\
D&Determiner gerund\\
pW{0,1,2}&wh-PP extraction\{position\}\\
W{0,1,2}&wh-NP extraction\{position\}\\
X&ECM (eXceptional case marking)\\
\end{tabular}
\end{description}

\noindent Numbers are assigned according to the position of the argument in the
declarative tree, as follows:

\begin{description}
\item\begin{tabular}{ll}
0&subject position\\
1&first argument (e.g. direct object)\\
2&second argument (e.g. indirect object)\\
\end{tabular}
\end{description}

\noindent The body of the name consists of a string of the following 
components, which corresponds to the leaves of the tree.  The anchor(s) of the
trees is(are) indicated by capitalizing the part of speech corresponding to the
anchor.

\begin{description}
\item\begin{tabular}{ll}
s&sentence\\
a&adjective\\
arb&adverb\\
be&{\it be}\\
c&relative complementizer\\
x&phrasal category\\
d&determiner\\
v&verb\\
lv&light verb\\
conj&conjunction\\
comp&complementizer\\
it&{\it it}\\
n&noun\\
p&preposition\\
to&to\\
pl&particle\\
by&{\it by}\\
neg&negation\\
\end{tabular}
\end{description}

\noindent As an example, the transitive declarative tree consists of a subject
NP, followed by a verb (which is the anchor), followed by the object NP.  This
translates into $\alpha$nx0Vnx1.  If the subject NP had been extracted, then
the tree would be $\alpha$W0nx0Vnx1.  A passive tree with the {\it by} phrase
in the same tree family would be $\alpha$nx1Vbynx0.  Note that even though the
object NP has moved to the subject position, it retains the object encoding
(nx1). 

\section{Assorted Initial Trees}

Trees that are not part of the tree families are generally gathered into
several files for convenience.  The various initial trees are located in {\tt
lex.trees}.  All the trees in this file should begin with an $\alpha$,
indicating that they are initial trees.  This is followed by the root category
which follows the naming conventions in the previous section (e.g. n for noun,
x for phrasal category).  The root category is in all capital letters.  After
the root category, the node leaves are named, beginning from the left, with the
anchor of the tree also being capitalized.  As an example, the $\alpha$NXN
tree is rooted by an NP node (NX) and anchored by a noun (N).

\section{Assorted Auxiliary Trees}

The auxiliary trees are mostly located in the buffers {\tt
prepositions.trees}, {\tt conjunctions.trees}, {\tt
determiners.trees}, {\tt advs-adjs.trees}, and {\tt modifiers.trees},
although a couple of other files also contain auxiliary trees.  The
auxiliary trees follow a slightly different naming convention from the
initial trees.  Since the root and foot nodes must be the same for the
auxiliary trees, the root nodes are not explicitly mentioned in the
names of auxiliary trees.  The trees are named according to the leaf
nodes, starting from the left, and capitalizing the anchor node.  All
auxiliary trees begin with a $\beta$, of course.  For example,
$\beta$ARBs, indicates a tree anchored by an adverb (ARB), that
adjoins onto the left of an S node (Note that S must be the foot node,
and therefore also the root node).

\subsection{Relative Clause Trees}
For relative clause trees, the following naming conventions have been
adopted: if the {\em wh}-moved NP is overt, it is not explicitly
represented. Instead the index of the site of movement
(0 for subject, 1 for object, 2 for indirect object) is appended to the
N. So $\beta$N0nx0Vnx1 is a subject
extraction relative clause with {\bf NP$_{w}$} substitution
and $\beta$N1nx0Vnx1 is an object extraction
relative clause. If the {\em wh}-moved NP is covert and Comp substitutes
in, the Comp node is represented by {\em c} in the tree name and the
index of the extraction site follows {\em c}. Thus
$\beta$Nc0nx0Vnx1 is a subject extraction
relative clause with Comp substitution. Adjunct trees are similar, except
that since the extracted material is not co-indexed to a trace, no index
is specified (cf. $\beta$Npxnx0Vnx1, which is an adjunct relative clause with
PP pied-piping, and $\beta$Ncnx0Vnx1, which is an adjunct relative clause
with Comp substitution). Cases of pied-piping, in which the pied-piped
material is part of the anchor have the anchor capitalized or spelled-out
(cf. $\beta$Nbynx0nx1Vbynx0 which is a relative clause with {\em by}-phrase
pied-piping and {\bf NP$_{w}$} substitution.).

