\appendix 
\chapter{Future Work} 
\label{future-work} 
 
\section{Adjective ordering} 
 
At this point, the treatment of adjectives in the XTAG English grammar does not 
include selectional or ordering restrictions.\footnote{This section is a repeat of information found in section~\ref{adj-modifier}.} Consequently, any 
adjective can adjoin onto any noun and on top of any other adjective already 
modifying a noun. All of the modified noun phrases shown in (\ref{ex:261})-(\ref{ex:264}) 
currently parse. 
 
\beginsentences
\sitem{big green bugs}\label{ex:261} 
\sitem{big green ideas}\label{ex:262} 
\sitem{colorless green ideas}\label{ex:263} 
\sitem{$\ast$green big ideas}\label{ex:264} 
\endsentences

 
While (\ref{ex:262})-(\ref{ex:264}) are all semantically anomalous, (\ref{ex:264}) also suffers 
from an ordering problem that makes it seem ungrammatical as well.  Since the 
XTAG grammar focuses on syntactic constructions, it should accept 
(\ref{ex:261})-(\ref{ex:263}) but not (\ref{ex:264}).  Both the auxiliary and determiner 
ordering systems are structured on the idea that certain types of lexical items 
(specified by features) can adjoin onto some types of lexical items, but not 
others.  We believe that an analysis of adjectival ordering would follow the 
same type of mechanism. 
 
\section{More work on Determiners} 
 
In addition to the analysis described in Chapter~\ref{det-comparitives}, there 
remains work to be done to complete the analysis of determiner constructions in 
English.\footnote{This section is from \cite{ircs:det98}.}  Although 
constructions such as determiner coordination are easily handled if 
overgeneration is allowed, blocking sequences such as {\it one and some} while 
allowing sequences such as {\it five or ten} still remains to be worked out. 
There are still a handful of determiners that are not currently handled by our 
system.  We do not have an analysis to handle {\it most}, {\it such}, {\it certain}, {\it other} and {\it own}\footnote{The behavior of {\it own} is sufficiently unlike other determiners that it most likely needs a tree of its own, adjoining onto the right-hand side of genitive determiners.}.  In 
addition, there is a set of lexical items that we consider adjectives ({\it enough}, {\it less}, {\it more} and {\it much}) that have the property that 
they cannot cooccur with determiners.  We feel that a complete analysis of 
determiners should be able to account for this phenomenon, as well. 
 
%\section{{\it -ing} adjectives} 
 
%An analysis has already been provided for past participal ({\it -ed}) 
%adjectives (as in sentence~ (\ref{ex:265})), which are restricted to the 
%Transitive Verb family.\footnote{This analysis may need to be extended %to the Transitive Verb particle family as well.}  A similar analysis 
%needs to take place for the present participle~({\it -ing}) used as a 
%pre-nominal modifier.  This type of adjective, however, does not seem 
%to be as restricted as the~{\it -ed} adjectives, since verbs in other 
%tree families seem to exhibit this alternation as well 
%(e.g. sentences~(\ref{ex:266}) and (\ref{ex:267})). 
 
%\enumsentence{The murdered man was a doctoral student at UPenn .} 
%\enumsentence{The man died .} 
%\enumsentence{The dying man pleaded for his life .} 
 
 
\section{Verb selectional restrictions} 
 
Although we explicitly do not want to model semantics in the XTAG grammar, 
there is some work along the syntax/semantics interface that would help reduce 
syntactic ambiguity and thus decrease the number of semantically anomalous 
parses.  In particular, verb selectional restrictions, particularly for PP 
arguments and adjuncts, would be quite useful.  With the exception of the 
required {\it to} in the Ditransitive with PP Shift tree family (Tnx0Vnx1Pnx2), 
any preposition is allowed in the tree families that have prepositions as their 
arguments.  In addition, there are no restrictions as to which prepositions are 
allowed to adjoin onto a given verb.  The sentences in (\ref{ex:265})-(\ref{ex:267}) are 
all currently accepted by the XTAG grammar.  Their violations are stronger than 
would be expected from purely semantic violations, however, and the presence of 
verb selectional restrictions on PP's would keep these sentences from being 
accepted. 
 
\beginsentences
\sitem{\#survivors walked of the street .}\label{ex:265} 
\sitem{\#The man about the earthquake survived .}\label{ex:266} 
\sitem{\#The president arranged on a meeting .}\label{ex:267} 
\endsentences

 
\section{Thematic Roles} 
 
Elementary trees in TAGs capture several notions of locality, with the most 
primary of these being locality of $\theta$-role assignment.  Each elementary 
tree has associated with it the $\theta$-roles assigned by the anchor of that 
elementary tree.  In the current XTAG system, while the notion of locality of 
$\theta$-role assignment within an elementary tree has been implicit, the 
$\theta$-roles assigned by a head have not been explicitly represented in the 
elementary tree. Incorporating $\theta$-role information will make the 
elementary trees more informative and will enable efficient pruning of spurious 
derivations when embedded into a specific context.  In the case of a 
Synchronous TAG, $\theta$-roles can also be used to automatically establish 
links between two elementary trees, one in the object language and one in the 
target language. 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
 
