\chapter{Getting Around}

This technical report presents the English XTAG grammar as implemented by
the XTAG Research Group at the University of Pennsylvania.  The technical
report is organized into four parts, plus a set of appendices.  Part 1
contains general information about the XTAG system and some of the
underlying mechanisms that help shape the grammar.
Chapter~\ref{intro-FBLTAG} contains an introduction to the formalism behind
the grammar and parser, while Chapter~\ref{overview} contains information
about the entire XTAG system.  Linguists interested solely in the grammar
of the XTAG system may safely skip these chapters. Chapter~\ref{underview}
contains a description of the organization of the grammar as tree families
and subcategorization frames, and discusses some of the linguistic
principles that underlie the XTAG grammar, including the distinction
between complements and adjuncts, and how case is handled.

The actual description of the grammar begins with Part 2, and is contained
in the following three parts.  Parts 2 and 3 contains information on the
verb classes and the types of trees allowed within the verb classes,
respectively, while Part 4 contains information on trees not included in
the verb classes (e.g.  NP's, PP's, various modifiers, etc).
Chapter~\ref{table-intro} of Part 2 contains a table that attempts to
provide an overview of the verb classes and tree types by providing a
graphical indication of which tree types are allowed in which verb classes.
This has been cross-indexed to tree figures shown in the tech report.
Chapter~\ref{verb-classes} contains an overview of all of the verb classes
in the XTAG grammar.  The rest of Part 2 contains more details on several
of the more interesting verb classes, including ergatives, sentential
subjects, sentential complements, copula, raising verbs, small clauses,
ditransitives, ditransitives, datives, verbs selecting PP complements,
resultatives and it-clefts.

Part 3 contains information on some of the tree types that are available within
the verb classes.  These tree types correspond to what would be transformations
in a movement based approach.  Not all of these types of trees are contained in
all of the verb classes.  The table (previously mentioned) in Part 2 contains a
list of the tree types and indicates which verb classes each occurs in.  

Part 4 focuses on the non-verb class trees in the grammar.  NP's and
determiners are presented in Chapter~\ref{det-comparitives}, while the
various modifier trees are presented in Chapter~\ref{modifiers}.  Auxiliary
verbs, which are classed separate from the verb classes, are presented in
Chapter~\ref{auxiliaries}, while certain types of conjunctions are shown in
Chapter~\ref{conjunction}.  The XTAG treatment of comparatives is presented
in Chapter~\ref{compars-chapter}, and our treatment of punctuation is
discussed in Chapter~\ref{punct-chapt}.

Throughout the technical report, mention is occasionally made of changes or
analyses that we hope to incorporate in the future.
Appendix~\ref{future-work} details a list of these and other future work.
The appendices also contain information on some of the nitty gritty details
of the XTAG grammar, including a system of metarules which can be used for
grammar development and maintenance in Appendix~\ref{metarules}. A system
for the organization of the grammar in terms of an inheritance hierarchy is
in Appendix~\ref{lexorg}. The tree naming conventions used in XTAG are
explained in detail in Appendix~\ref{tree-naming}, and a comprehensive list
of the features used in the grammar is given in Appendix~{\ref{features}.
Appendix~\ref{evaluation} contains an evaluation of the XTAG grammar,
including comparisons with other wide coverage grammars.


